---
title: "Assignment 2"
author: "Briana Flynn"
date: "9/22/2020"
output: html_document
---

# Instructions

Select a large area and create a map of the area, with its associated subareas, using each of three different map projections. Indicate the map projection you’ve used for each. Identify a variable and use at least three different methods to show how it varies among the subareas.

# Loading Data & Identifying Area of Interest

I began by loading a number of libraries and new packages. After loading a world map to R for the first time, I chose South America as the large area I would re-project, transform, and collect data for in this assignment. 

```{r}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(rgdal)
library(grid)
```

```{r}
sacountries <- ne_countries(continent = "South America", returnclass = "sf")

sa_map <-ggplot (sacountries) +
  geom_sf() +
  theme_bw()

sa_map
```

# Representing South America Using Alternate Map Projections

With South America as my chosen "large area" for Assignment 2, I looked for interesting and/or suitable map projections by searching for standard and then South America-focused projection systems. I settled on three projections:
  
  1. The Robinson Projection ("ROBINSON"): commonly used for a map projection of a world map which shows the entire  world at once, this projection is analagous to how South America might appear on a world map.
  2. South America Albers Equal Area Conic Projection ("ALBERS"): a modified version of the Albers Equal Conic Area projection, this projection uses two standard parallels to reduce some of the distortion of a projection with one standard parallel. 
  3. Cylindrical Equal Area Projection ("CEA"): meridians are mapped to equally spaces vertical lines and circles of latitude are mapped to horizontal lines.

```{r}
ROBINSON <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
ALBERS <- "+proj=lcc +lat_1=-5 +lat_2=-42 +lat_0=-32 +lon_0=-60 +x_0=0 +y_0=0 +ellps=aust_SA +units=m +no_defs"
CEA <- "+proj=cea +lon_0=-60 +lat_ts=-5 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"
```


## Robinson Projection

```{r}
ggplot(sacountries) +
  geom_sf() +
  coord_sf(crs = ROBINSON) +
  theme_bw()
```

## Albers Equal Area Conic Projection

```{r}
ggplot(sacountries) +
  geom_sf() +
  coord_sf(crs = ALBERS) +
  theme_bw()
```

## Cylndrical Equal Area Projection

```{r}
ggplot(sacountries) +
  geom_sf() +
  coord_sf(crs = CEA) +
  theme_bw()
```

# Transforming the Map and Joining External Data

Upon mapping South America with the three projections above, I transformed the map with the Albers projection for future mapping. I then pulled in data from the World Bank on the percentage of the population living in urban areas, 1960-2019. I joined the World Bank data with the transformed map. 

```{r}
SAcountries_transformed <- sacountries %>%
  st_transform(crs = ALBERS)
```

```{r}
urbanpop <- read_csv("UrbanPopDatav2.csv")
```

```{r}
sacountries_urbanpop <- SAcountries_transformed %>%
  left_join(urbanpop, by = c("name" = "CountryName")) %>%
  select(name, UrbanPop1960, UrbanPop2019)
```

```{r}
data.frame(sacountries_urbanpop)
```

# Representing Urban Population Variable for South American Countries

## Mapping Percentage Urban Population with Gradient Color

```{r}
 ggplot(sacountries_urbanpop, aes(fill = UrbanPop2019)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(option = "inferno", 
    name = "Percentage of\npopulation\nliving in\nurban\nareas 2019",
    breaks = seq(0, 100, by = 10),
    labels = formatC(seq(0, 100, by = 10), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9))
```

Note: Falkand Islands was excluded from the World Bank dataset.

## Mapping Percentage Urban Population with a Cartogram

```{r}
SAcountries_transformed = na.omit(SAcountries_transformed)
```

```{r}
sacountries_urbanpop = na.omit(sacountries_urbanpop)
```

```{r message=FALSE, warning=FALSE}
UrbanPopCartogramCont <- sacountries_urbanpop %>%
  cartogram_cont("UrbanPop2019")

ggplot(UrbanPopCartogramCont, aes(fill = UrbanPop2019)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(option = "inferno",
     name = "Percentage of\npopulation\nliving in\nurban\nareas 2019",
    breaks = seq(0, 100, by = 10),
    labels = formatC(seq(0, 100, by = 10), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9))
```

## Mapping Percentage Urban Population with a Non-continuous Cartogram

```{r message=FALSE, warning=FALSE}
urbanpop_cartogram_ncont <- sacountries_urbanpop %>%
  cartogram_ncont("UrbanPop2019")

ggplot(urbanpop_cartogram_ncont, aes(fill = UrbanPop2019)) +
 geom_sf(color = NA) +
  scale_fill_viridis_c(option = "inferno",
     name = "Percentage of\npopulation\nliving in\nurban\nareas 2019",
    breaks = seq(0, 100, by = 10),
    labels = formatC(seq(0, 100, by = 10), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9))
```
