---
title: "Assignment 3"
author: "Briana Flynn"
date: "9/25/2020"
output: html_document
---

# Points and Polygons 

```{r message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

```{r}
citycouncildistricts <- st_read( "http://bostonopendata-boston.opendata.arcgis.com/datasets/29c9c2e4ad884b299bc9b0ea571269bc_0.kml",
                  quiet = TRUE) 

water <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2b3c0fa13b1c468eb702a3645fcc6bcb_5.kml", 
                 quiet = TRUE)

publiclibraries <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/cb00f9248aa6404ab741071ca3806c0e_6.kml", 
                   quiet = TRUE) 

universities <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/cbf14bb032ef4bd38e20429f71acb61a_2.kml", 
                 quiet = TRUE)

mainstreetdistricts <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/440c7ec0178d4c8593aecef7ea96bb4d_0.kml",
                               quiet = TRUE)
```

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

citycouncildistricts <- citycouncildistricts %>%
  st_transform(MA_state_plane)

water <- water %>%
  st_transform(MA_state_plane)

publiclibraries <- publiclibraries %>%
  st_transform(MA_state_plane)

universities <- universities %>%
  st_transform(MA_state_plane)

mainstreetdistricts <- mainstreetdistricts %>%
  st_transform(MA_state_plane)
```

# Base Figure for Context

```{r}
ggplot(water) +
  geom_sf(fill = "darkblue", color = NA) +
  geom_sf(data = publiclibraries, fill = NA, color = "purple") +
  geom_sf(data = universities, fill = NA, color = "orange") +
  geom_sf(data= citycouncildistricts, fill = NA, color = "black") +
  theme_map() +
  annotation_scale()
```

# Figure 1: The number and percentage of public libraries within 500 meters of a college or university

```{r include=FALSE}
university_buffer <- st_buffer(universities, dist = 500) %>%
  st_union()

ggplot(university_buffer) +
  geom_sf() +
  theme_map()
```

```{r include=FALSE}
libraries_universities <- publiclibraries[university_buffer,]
  
ggplot(university_buffer) +
  geom_sf() +
  geom_sf(data = libraries_universities, 
          color = "purple", 
          size = 2) +
  theme_map()
```

```{r warning=FALSE, include=FALSE}
publiclibraries <- publiclibraries %>%
  st_join(libraries_universities) %>%
  mutate(by_universities = !is.na(Name.y))
```

```{r}
n_publiclibraries_universities <- sum(publiclibraries$by_universities)

n_publiclibraries_universities
```
```{r}
n_publiclibraries <- length(publiclibraries$by_universities)

pct_publiclibraries_universities <- n_publiclibraries_universities / n_publiclibraries

pct_publiclibraries_universities
```

```{r}
left_side  <- st_bbox(citycouncildistricts)$xmin
top_side <- st_bbox(citycouncildistricts)$ymax

ggplot(water) +
  geom_sf(fill = "darkblue", color = NA) +
  geom_sf(data = publiclibraries, size = 2,
          aes(color = by_universities)) +
  geom_sf(data= citycouncildistricts, fill = NA, color = "black") +
  scale_color_manual(values = c("gray", "purple"),
          name = "Public libraries\nby distance to a college or university", 
          labels = c("No college or university within 500 m",
                     "College or university within 30 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_publiclibraries, big.mark = ","),
                         " public libraries\nin Boston ", 
                         prettyNum(n_publiclibraries_universities, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_publiclibraries_universities, digits = 0),
                         "%) are\nwithin 500 meters\nof a university.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

There are 26 public libraries (23% of all public libraries) within 500 meters of a college or university. 

# Figure 2: Average number of public libraries by City Council District

```{r}
citycouncildistricts <- citycouncildistricts %>%
  mutate(num_publiclibraries = lengths(st_covers(citycouncildistricts, publiclibraries)))

ggplot(citycouncildistricts) +
  geom_sf(color = NA, 
          aes(fill = num_publiclibraries)) +
  scale_fill_viridis_c(name = "Boston council districts\nby number of public libraries",
                       breaks = breaks <- seq(0, 10, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "public libraries")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

```{r}
show(citycouncildistricts)
```

From reading the new "lengths" column in the dataframe, I know that the districts range from having two libraries to four libraries, and that the average number of public libraries per district is 2.88.

# Figure 3: Number of colleges or unviersities by City Council District

```{r}
citycouncildistricts <- citycouncildistricts %>%
  mutate(num_universities = lengths(st_covers(citycouncildistricts, universities)))

ggplot(citycouncildistricts) +
  geom_sf(color = NA, 
          aes(fill = num_universities)) +
  scale_fill_viridis_c(name = "Boston council districts\nby number of colleges and universities",
                       breaks = breaks <- seq(0, 40, by = 10),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "colleges and universities")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

```{r}
show(citycouncildistricts)
```

From reading the new "lengths" column in the dataframe, I know that the districts range from having zero colleges or universities to having 31 colleges or universities, with the average being 6.44 per district. 


# FIgure 4. Colleges and universities by distance to the nearest public library

```{r include=FALSE}
universities <- universities %>%
  mutate(publiclibraries_dist = st_nn(universities, publiclibraries, 
                           returnDist = TRUE)$dist) %>%
  mutate(publiclibraries_dist = as.numeric(publiclibraries_dist))
```
```{r}
avg_publiclibraries_dist <- mean(universities$publiclibraries_dist)

avg_publiclibraries_dist
```

```{r}
right_side <- st_bbox(citycouncildistricts)$xmax
left_side  <- st_bbox(citycouncildistricts)$xmin
top_side <- st_bbox(citycouncildistricts)$ymax
bottom_side <- st_bbox(citycouncildistricts)$ymin


ggplot(water) +
  geom_sf(fill = "darkblue", color = NA) +
  geom_sf(data = universities, size = 2,
          aes(color = publiclibraries_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "Colleges and unviersities\nby distance to\na public libary") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 800, 
           y = top_side - 1600, 
           label = paste("On average, a college or university\nis ", 
                         prettyNum(avg_publiclibraries_dist, digits = 3),
                         " meters from a public library.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

Boston's colleges and universities are an average of 937 meters from a public library. 

# Figure 5. Boston City Council Districts intersecting with a Main Street District

```{r include=FALSE}
citycouncildistricts <- citycouncildistricts %>%
  mutate(num_mainstreetdistricts = lengths(st_intersects(citycouncildistricts, mainstreetdistricts))) %>%
  mutate(has_mainstreetdistricts = num_mainstreetdistricts > 0)
```

```{r}
n_mainstreet_citycouncildistricts <- sum(citycouncildistricts$has_mainstreetdistricts)

n_mainstreet_citycouncildistricts
```
```{r}
ggplot(mainstreetdistricts) +
  geom_sf(fill = "pink", color = NA) +
  geom_sf(data = publiclibraries, fill = NA, color = "purple") +
  geom_sf(data = universities, fill = NA, color = "orange") +
  geom_sf(data= citycouncildistricts, fill = NA, color = "black") +
  theme_map() +
  annotation_scale()
```

```{r}
left_side  <- st_bbox(citycouncildistricts)$xmin
top_side <- st_bbox(citycouncildistricts)$ymax

ggplot(mainstreetdistricts) +
  geom_sf(fill = "black", color = NA) +
  geom_sf(data = citycouncildistricts,
          aes(fill = has_mainstreetdistricts)) +
  scale_fill_manual(values = c("white", "yellow"),
          name = "Boston council districts\nby presence of a main street district", 
          labels = c("Council district without\na main street district",
                     "Council district with a\nmain street district")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste(n_mainstreet_citycouncildistricts,
                         "of Boston's", 
                         length(citycouncildistricts$Name),
                         "council districts contain\nor overlap with", 
                         "a main street district."),
           hjust = 0, vjust = 0, size = 2) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

Nine council districts (100%) contain a main street district.

# Figure 6. Main Street Districts overlapping with a body of water

```{r include=FALSE}
mainstreetdistricts <- mainstreetdistricts %>%
  mutate(num_water = lengths(st_overlaps(mainstreetdistricts, water))) %>%
  mutate(has_water = num_water > 0)
```

```{r}
n_water_mainstreetdistricts <- sum(mainstreetdistricts$has_water)

n_water_mainstreetdistricts
```

```{r}
left_side  <- st_bbox(water)$xmin
top_side <- st_bbox(water)$ymax

ggplot(water) +
  geom_sf(fill = "darkblue", color = NA) +
  geom_sf(data = mainstreetdistricts,
          aes(fill = has_water)) +
  scale_fill_manual(values = c("white", "yellow"),
          name = "Boston Main Street Districts\nby presence of a body of water", 
          labels = c("Main Street Districts without\nan overlapping body of water",
                     "Main Street Districts with an\noverlapping body of water")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 1000, 
           label = paste(n_water_mainstreetdistricts ,
                         "of Boston's", 
                         length(mainstreetdistricts$Name),
                         "main street districts contain\nor overlap with", 
                         "a body of water."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
Zero main street districts (0%) overlap with a body of water. 


